MS Depression Paper, post replication, started 2022 09 21

### 2022-09-29 msdepression/scripts/MS_Depression_First_Paper_started_20220728/msdepression_n380_started_2022092022_post_replication.Rmd
visualizations

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(dplyr)
library(visreg)
library(lme4)
library(lmerTest)
library(ggplot2)
library(reshape2)
library(pwr)
library(mgcv)
library(boot)
library(boot.pval)
library(lsr)
library(sjPlot)
library(coin)
library(rstatix)
library(effectsize)

@@

### Interaction Network x Diagnosis
#scale=1

@@

@@

violin_dep_vs_nondep_scaled_by_network_size <- ggplot(df_proportion_network_volume_and_diagnosis, aes(x=depGroupVar, y=proportion_volume_lost_per_total_network_size_sum, fill=depGroupVar)) +
  geom_violin() +  
  geom_boxplot(width=0.1, color="black", fill="white", outlier.size = 0.2) + 
  scale_x_discrete(labels=c("1" = "NonDepressed", "2" = "Depressed")) + 
  ggtitle(paste0("Proportion Fascicle Volume Affected by Diagnosis")) +
  ylab("Proportion (sum total volume lost/Net Vol) ") +
  ylim(0,1) + 
  scale_fill_manual(values=c("#AB19E6", "#1954E6")) + 
  theme(plot.title = element_text(size=16, hjust = 0.55),
        axis.text = element_text(size = 14, colour = "black"), axis.title.x = element_blank(), 
        axis.title.y = element_text(size=16),
        axis.line = element_line(size = 1.5, colour = "black", linetype=1),
        legend.position = "none",
        aspect.ratio=7/5,
        panel.background = element_rect(colour = "white", fill = "white"))
print(violin_dep_vs_nondep_scaled_by_network_size) #

@@

#interaction

diagnosis_by_network_interaction_lmer_plot <- plot_model(diagnosis_by_network_interaction_lmer, 
                                                         type = "int",
                                                         axis.lim = c(0,1),
                                                         colors = c("#ff00ff", "#3df2f2"),
                                                         line.size = 2,
                                                         dodge=0.2,
                                                         axis.title = c("Diagnosis", "Proportion Affected"),
                                                         title = "Diagnosis x Depression Network Disease") +

  theme(axis.text.x = element_text(size=16, color="black"),
        axis.text.y = element_text(size=16, color="black"),
        axis.title.x = element_text(size =16), 
        axis.title.y = element_text(size =16), 
        title = element_text(size=18),
        axis.line = element_line(size = 1.5, colour = "black", linetype=1),
        aspect.ratio = 5/3,
        panel.background = element_rect(colour = "white", fill = "white"))

print(diagnosis_by_network_interaction_lmer_plot)

#save as pdf
pdf(file=paste0(homedir, "/results/Figures_for_msdepression_paper/diagnosis_by_network_interaction_lmer_range0to1.pdf"))
  diagnosis_by_network_interaction_lmer_plot
dev.off()

```
### Effect size regression (effect size x overlap with dep net)
``` {r effect_size_regression}
dep_and_healthy_groups_for_ICD_analysis <- df_demo_and_fascicles_prop_and_vol %>% 
  filter(depGroupVar != 0)  %>% 
 # mutate(total_fascicle_vol_lost = rowMeans(across(all_of(fascicle_names_w_vol_suffix)))) %>%
  group_by(EMPI) %>% 
  arrange(EXAM_DATE) %>% 
  slice(1) %>% 
  ungroup()

>>>>

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(dplyr)
library(visreg)
library(lme4)
library(lmerTest)
library(ggplot2)
library(reshape2)
library(pwr)
library(mgcv)
library(boot)
library(boot.pval)
library(lsr)
library(sjPlot)
library(coin)
library(rstatix)
library(effectsize)
library(colormod)
library(ggpubr)

@@

### Interaction Network x Diagnosis

@@

#### color scheme
Diagnosis_colors <- c("#AB19E6", "#1954E6")
Diagnosis_colors_faded <- as.vector(sapply(Diagnosis_colors, function(color) {decrease_saturation(color, 0.3)})) 
Network_colors <- c("#ff00ff", "#3df2f2")
Network_colors_faded <- as.vector(sapply(Network_colors, function(color) {decrease_saturation(color, 0.3)}))

#label scheme
Network_labels <- c("Depression Network", "Nondepression Network")
Network_labels_wrapped <- c("Depression\nNetwork", "Nondepression\n Network")
Diagnosis_labels <- c("Depressed", "Nondepressed")

@@

violin_dep_vs_nondep_scaled_by_network_size <- ggplot(df_proportion_network_volume_and_diagnosis, aes(x=depGroupVar, y=proportion_volume_lost_per_total_network_size_sum, fill=depGroupVar)) +
  geom_violin() +  
  geom_boxplot(width=0.1, color="black", fill="white", outlier.size = 0.2) + 
  scale_x_discrete(labels=c("1" = "NonDepressed", "2" = "Depressed")) + 
  ggtitle(paste0("Proportion Fascicle Volume Affected by Diagnosis")) +
  ylab("Proportion (sum total volume lost/Net Vol) ") +
  ylim(0,1) + 
  scale_fill_manual(values=Diagnosis_colors) + 
  theme(plot.title = element_text(size=16, hjust = 0.55),
        axis.text = element_text(size = 14, colour = "black"), axis.title.x = element_blank(), 
        axis.title.y = element_text(size=16),
        axis.line = element_line(size = 1.5, colour = "black", linetype=1),
        legend.position = "none",
        aspect.ratio=7/5,
        panel.background = element_rect(colour = "white", fill = "white"))
print(violin_dep_vs_nondep_scaled_by_network_size) #

@@

# bar graph instead of violin

df_proportion_network_volume_and_diagnosis_summary <- df_proportion_network_volume_and_diagnosis %>%
  group_by(depGroupVar) %>%
  summarise(
    mean = mean(proportion_volume_lost_per_total_network_size_sum),
    sd = sd(proportion_volume_lost_per_total_network_size_sum),
    n = n(),
    se = sd / sqrt(n)
  )

bar_dep_vs_nondep_scaled_by_network_size <- ggplot(df_proportion_network_volume_and_diagnosis_summary, aes(x=depGroupVar, y=mean, fill=depGroupVar)) +
  geom_bar(stat = "identity", fill=Diagnosis_colors_faded, color = Diagnosis_colors, size = 1.5, width = 0.8) +  
  geom_errorbar(aes(ymin=mean-se,ymax=mean+se), width=0.2, size=1.5, color = 'black') + #, position=position_dodge(width=0.5), width = 0.2,size = 1.5, color='black') +
  scale_x_discrete(labels=c("Non-depressed" = "Nondepressed", "Depressed" = "Depressed")) + 
  ggtitle(paste0("Proportion Fascicle Volume Affected by Diagnosis")) +
  ylab("Proportion (sum total volume lost/Net Vol) ") +
  ylim(0,0.5) + 
  theme(plot.title = element_text(size=16, hjust = 0.55),
        axis.text = element_text(size = 16, colour = "black"), axis.title.x = element_blank(), 
        axis.title.y = element_text(size=16),
        axis.line = element_line(size = 1.5, colour = "black", linetype=1),
        legend.position = "none",
        aspect.ratio=7/5,
        panel.background = element_rect(colour = "white", fill = "white"))
print(bar_dep_vs_nondep_scaled_by_network_size) #

#save pdf
pdf(file=paste0(homedir, "/results/Figures_for_msdepression_paper/bar_dep_vs_nondep_scaled_by_network_size.pdf"))
  bar_dep_vs_nondep_scaled_by_network_size
dev.off()

#plot interaction with ggplot rather than plot_model

df_interaction <- expand.grid(Network = levels(df_empi_dx_network_proportions$Network), Diagnosis = levels(df_empi_dx_network_proportions$Diagnosis), subject = "8450972724") # just pick a random person

model_data <- get_model_data(model = diagnosis_by_network_interaction_lmer, type = "int")

df_interaction_fit_upper_lower <- cbind(df_interaction, model_data$predicted, model_data$conf.high, model_data$conf.low)

### line

interaction_line <- ggplot(data = df_interaction_fit_upper_lower, aes(x = Network, y = model_data$predicted, color = Diagnosis)) +
  geom_line(aes(group = Diagnosis), size = 1.5) +
  scale_color_manual(values=Diagnosis_colors, labels=Diagnosis_labels)+ 
  geom_point(size = 3) + 
 # geom_errorbar(aes(ymin=model_data$conf.low,ymax=model_data$conf.high), width = 0.2,size = 1, color='black') +
  scale_x_discrete(labels=Network_labels) +
  ylim(0,0.5) + 
    xlab("Diagnosis") + 
    ylab("Proportion Affected") + 
    theme(axis.text.x = element_text(size=16, color="black"),
          axis.text.y = element_text(size=16, color="black"),
          axis.title.x = element_blank(), #element_text(size =16), 
          axis.title.y = element_text(size =16), 
          title = element_text(size=18),
          axis.line = element_line(size = 1.5, colour = "black", linetype=1),
          aspect.ratio = 7/5,
          panel.background = element_rect(colour = "white", fill = "white"),
          legend.key=element_blank(),
          legend.title = element_blank(),
          legend.text = element_text(size = 14), legend.position = "top")

#save as pdf
pdf(file=paste0(homedir, "/results/Figures_for_msdepression_paper/diagnosis_by_network_interaction_line_graph_range0to05_point.pdf"))
  interaction_line
dev.off()

#geom_point
interaction_ggplot_point <- ggplot(data = df_interaction_fit_upper_lower, mapping = aes(x = Diagnosis, y = model_data$predicted, color = Network, group = interaction(Diagnosis, Network))) + 
    geom_errorbar(aes(ymin=model_data$conf.low,ymax=model_data$conf.high), position=position_dodge(width=0.5), width = 0.2,size = 1.5, color='black') +
    geom_point(position = position_dodge(width=0.5), size = 5) + 
    scale_color_manual(values=Network_colors, labels = c("Depression Network", "Nondepression Network")) + 
    ylim(0,1) + 
    xlab("Diagnosis") + 
    ylab("Proportion Affected") + 
    theme(axis.text.x = element_text(size=16, color="black"),
          axis.text.y = element_text(size=16, color="black"),
          axis.title.x = element_blank(), #element_text(size =16), 
          axis.title.y = element_text(size =16), 
          title = element_text(size=18),
          axis.line = element_line(size = 1.5, colour = "black", linetype=1),
          aspect.ratio = 5/3,
          panel.background = element_rect(colour = "white", fill = "white"),
          legend.key=element_blank(),
          legend.title = element_blank(),
          legend.text = element_text(size = 14), #,
          legend.position = "top") #+ 
 #   ggtitle("Diagnosis x Depression Network Disease")
print(interaction_ggplot_point)

#save as pdf
pdf(file=paste0(homedir, "/results/Figures_for_msdepression_paper/diagnosis_by_network_interaction_plot_range0to1_point.pdf"))
  interaction_ggplot_point
dev.off()

#geom_point 0.2-0.5
interaction_ggplot_point_02_05 <- ggplot(data = df_interaction_fit_upper_lower, mapping = aes(x = Diagnosis, y = model_data$predicted, color = Network, group = interaction(Diagnosis, Network))) + 
    geom_errorbar(aes(ymin=model_data$conf.low,ymax=model_data$conf.high), position=position_dodge(width=0.5), width = 0.2,size = 1.5, color='black') +
    geom_point(position = position_dodge(width=0.5), size = 5) + 
    scale_color_manual(values=Network_colors, labels = c("Depression Network", "Nondepression Network")) + 
    ylim(0.2,0.5) + 
    xlab("Diagnosis") + 
    ylab("Proportion Affected") + 
    theme(axis.text.x = element_text(size=16, color="black"),
          axis.text.y = element_text(size=16, color="black"),
          axis.title.x = element_blank(), #element_text(size =16), 
          axis.title.y = element_text(size =16), 
          title = element_text(size=18),
          axis.line = element_line(size = 1.5, colour = "black", linetype=1),
          aspect.ratio = 5/3,
          panel.background = element_rect(colour = "white", fill = "white"),
          legend.key=element_blank(),
          legend.title = element_blank(),
          legend.text = element_text(size = 14), #,
          legend.position = "top") #+ 
 #   ggtitle("Diagnosis x Depression Network Disease")
print(interaction_ggplot_point_02_05)

#save as pdf
pdf(file=paste0(homedir, "/results/Figures_for_msdepression_paper/diagnosis_by_network_interaction_plot_range0_2to0_5_point.pdf"))
  interaction_ggplot_point_02_05
dev.off()

#geom_bar
interaction_ggplot_bar <- ggplot(data = df_interaction_fit_upper_lower, mapping = aes(x = Diagnosis, y = model_data$predicted, color = Network,  group = interaction(Diagnosis, Network))) + 
    geom_col(position = position_dodge(width=0.6), fill = c(Network_colors_faded, Network_colors_faded), color = c(Network_colors, Network_colors), width = 0.5, size = 1.5) + 
    geom_errorbar(aes(ymin=model_data$conf.low,ymax=model_data$conf.high), position=position_dodge(width=0.6), width = 0.2,size = 1.5, color='black') +
    #geom_point(position = position_dodge(width=0.5), size = 5) + 
    scale_color_manual(values=Network_colors) + #, labels = c("Depression Network", "Nondepression Network"))+ 
    ylim(0,0.5) + 
    xlab("Diagnosis") + 
    ylab("Proportion Affected") + 
    theme(axis.text.x = element_text(size=16, color="black"),
          axis.text.y = element_text(size=16, color="black"),
          axis.title.x = element_blank(), #element_text(size =16), 
          axis.title.y = element_text(size =16), 
          plot.title = element_text(size=16, hjust = 0.5),
          axis.line = element_line(size = 1.5, colour = "black", linetype=1),
          aspect.ratio = 7/5,
          panel.background = element_rect(colour = "white", fill = "white"),
          legend.key=element_blank(),
          legend.title = element_blank(),
          legend.text = element_text(size = 14),
          legend.position = "top")+

    ggtitle("Diagnosis x Depression Network Disease")
print(interaction_ggplot_bar)

#save as pdf
pdf(file=paste0(homedir, "/results/Figures_for_msdepression_paper/diagnosis_by_network_interaction_plot_range0to05_bar.pdf"))
  interaction_ggplot_bar
dev.off()

### wrap

diagnosis_and_interaction_bar_combined <- ggarrange(bar_dep_vs_nondep_scaled_by_network_size,interaction_ggplot_bar, 
                                                labels = c("A", "B"),
                                                ncol = 2, nrow = 1)

pdf(file=paste0(homedir, "/results/Figures_for_msdepression_paper/diagnosis_intx_line_and_intx_bargraph_range0to05.pdf"))
  diagnosis_and_interaction_bar_combined

diagnosis_and_interaction_all_combined <- ggarrange(bar_dep_vs_nondep_scaled_by_network_size, interaction_line,interaction_ggplot_bar, 
                                                labels = c("A", "B", "C"),
                                                ncol = 3, nrow = 1)

pdf(file=paste0(homedir, "/results/Figures_for_msdepression_paper/diagnosis_intx_line_and_intx_bargraph_range0to05.pdf"))
  diagnosis_and_interaction_all_combined
dev.off()
```

### Effect size regression (effect size x overlap with dep net)
``` {r effect_size_regression}
dep_and_healthy_groups_for_ICD_analysis <- df_demo_and_fascicles_prop_and_vol %>% 
  filter(depGroupVar != 0)  %>% 
 # mutate(total_fascicle_vol_lost = rowMeans(across(all_of(fascicle_names_w_vol_suffix)))) %>%
  group_by(EMPI) %>% 
  arrange(EXAM_DATE) %>% 
  slice(1) %>% 
  ungroup()
